---
title: "Downscaling Workflow"
author: "Maite"
date: "15 8 2021"
output: github_document
---

```{r setup, eval=F, include=F}
knitr::opts_chunk$set(echo = TRUE)
```

# This document 
This document summarizes the different steps required for the downscaling approach described in XXX and links to the scripts within this repository in the adequate order. The resulting downscaling model and necessary predictor data to downscale a MODIS scene captured during the austral summer months during daylight conditions and the McMurdo Dry Valleys can be installed from [Github repo downscaleLST.MDV](https://github.com/MLezamaValdes/downscaleLST.MDV)  


## Basic settings 
- load libraries
- define time range 
- create directories for Landsat and MODIS data to be downloaded
- read Area of Interest (AOI) shape
- define and source some functions


```{r, include=FALSE, eval=F}
# newarea can be 0 (no) or 1 (yes). If area is not new, it is assumed that the land outline shapefile and cropped rock outcrop is already available and doesn't need to be unzipped etc.
newarea <- 0
areaname <- "MDV"

# L8: either "Bt" or "L1"
L8downloadtype <- "Bt"


## time range parameters 
year <- c(2020:2013)
month <- c("01","02","03","04", "09", "10","11", "12")

# maximum time between satellite scenes 
timethres <- 0.6

scriptpath <- "C:/Users/mleza/OneDrive/Documents/PhD/work_packages/auto_downscaling_30m/downscale_controlscripts/data_prep/"
scriptpath_organized <- "C:/Users/mleza/OneDrive/Documents/PhD/work_packages/auto_downscaling_30m/downscale_controlscripts/data_prep/downscaling_organized/"

maindir <- "D:/downscaling_after_talk/" # this must point to an existing directory, rest is generated in setup
main <- paste0(maindir, "data_download_preprocessing/")
L8datpath <- paste0(main, "L8/")
modispath <- paste0(main, "MODIS/")
tdpath <-paste0(main, "timediff/")
cddir <- paste0(maindir, "clean_data/")
figurepath <- "C:/Users/mleza/OneDrive/Documents/PhD/work_packages/auto_downscaling_30m/paper/paper_draft/figures/"


##### set path to DEM, AOI, land outline

dempath <- "E:/new_downscaling/tiles_westcoast/" # this must point to an existing directory with dem inside
aoipath <-  "E:/new_downscaling/aoi/" # this must point to an existing directory with "Levy_MDV_actually.shp"
aoip <- list.files(aoipath, pattern="actually.shp", full.names = T)
clpath <- "E:/new_downscaling/coastline/Coastline_high_res_polygon/"

`%notin%` <- Negate(`%in%`)

## PATH TO TOOLS CMD

##### set paths for translating to SAGA
path_saga_norm <- "C:/OSGeo4W64/apps/saga-ltr/"
sagaCmd <- paste0(path_saga_norm, "saga_cmd.exe")
saga_outpath <- paste0(main, "SAGA_run/")

##### set paths for batch processing in HEG tool
# make an batchindir directory with the hdf files to batch convert and an batchoutdir directory with 
# the prm template file, where output files will be written to
# batchrunpath needs to be where MyHEG_batchScript.bat is located
batchrunpath <- "C:/Users/mleza/HEG/HEG_Win/bin/BatchRunning/BatchRunning/"
batchindir <- paste0(batchrunpath, "indir/") 
batchoutdir <- paste0(batchrunpath, "outdir/")
# for stacking images per month


source(paste0(scriptpath_organized, "0a_setup.R"))
#file.edit(paste0(scriptpath_organized, "0a_setup.R"))

# this is one raster with a complete coverage of the research area to use as a template 
template <- raster("E:/new_downscaling/clean_data/template_new.tif")

```

## 1 Preparing the Digital Elevation Model (DEM)
```{r, eval=F}
file.edit("1_DEM.R")
```

## 2 Selecting, downloading and processing Landsat Band 10 to Land Surface Temperature (LST)
Scenes that show little cloud contamination, are overlapping the AOI sufficiently and are temporally closely matching an available MODIS scene are being selected and queried.
```{r, eval=F}
file.edit(paste0(scriptpath_organized, "2_1_selectLandsat.R"))
file.edit(paste0(scriptpath_organized, "2_2a_downloadLandsat.R"))
file.edit(paste0(scriptpath_organized, "2_2b_downloadLandsat.R"))
file.edit(paste0(scriptpath_organized, "2_2c_place_scenes.R"))
file.edit(paste0(scriptpath_organized, "2_3_processLandsat.R"))
```

## 3 MODIS 
MOD11_L2 and MYD11_L2 data come already cloud masked by MODIS product MOD35_L2. 
```{r, eval=F}
file.edit("3_MODIS.R")
```


## 4 Creating a Landsat and MODIS data stack per month 
```{r, eval=F}
file.edit("4_make_sat_stack.R")
```

## 5 Create incidence angle and hillshading raster
For each unique MODIS scene capturing time, topographic solar shading and the solar incidence angle are calculated based on the DEM and the corresponding solar position. 
```{r, eval=F}
file.edit("Palma/5_ia_hs_Palma.R")
```


## 6 Match satellite LST data and solar insolation variables
```{r, eval=F}
file.edit("6_match_MOD_L_ia_hs.R")
```

## 7 extract data from raster stacks to table format
```{r, eval=F}
file.edit("Palma/7_stack_extraction_Palma.R")
```

