---
title: "Downscaling Workflow"
author: "Maite"
date: "15 8 2021"

output: 
 rmarkdown::html_document:
  #bookdown::html_document2:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 6
link-citations: yes
toc: yes
---

```{r setup, eval=F}
knitr::opts_chunk$set(echo = TRUE)
```

# This document 
This document summarizes the different steps required for the downscaling approach described in XXX and links to the scripts within this repository in the adequate order. 


## Basic settings 
- load libraries
- define time range 
- create directories for Landsat and MODIS data to be downloaded
- read Area of Interest (AOI) shape
- define and source some functions


```{r, include=FALSE, eval=F}
# newarea can be 0 (no) or 1 (yes). If area is not new, it is assumed that the land outline shapefile and cropped rock outcrop is already available and doesn't need to be unzipped etc.
newarea <- 0
areaname <- "MDV"

# L8: either "Bt" or "L1"
L8downloadtype <- "Bt"


## time range parameters 
year <- c(2020:2013)
month <- c("01","02","03","04", "09", "10","11", "12")

# maximum time between satellite scenes 
timethres <- 0.6

scriptpath <- "C:/Users/mleza/OneDrive/Documents/PhD/work_packages/auto_downscaling_30m/downscale_controlscripts/data_prep/"
scriptpath_organized <- "C:/Users/mleza/OneDrive/Documents/PhD/work_packages/auto_downscaling_30m/downscale_controlscripts/data_prep/downscaling_organized/"

maindir <- "D:/downscaling_after_talk/" # this must point to an existing directory, rest is generated in setup
main <- paste0(maindir, "data_download_preprocessing/")
L8datpath <- paste0(main, "L8/")
modispath <- paste0(main, "MODIS/")
tdpath <-paste0(main, "timediff/")
cddir <- paste0(maindir, "clean_data/")
figurepath <- "C:/Users/mleza/OneDrive/Documents/PhD/work_packages/auto_downscaling_30m/paper/paper_draft/figures/"


##### set path to DEM, AOI, land outline

dempath <- "E:/new_downscaling/tiles_westcoast/" # this must point to an existing directory with dem inside
aoipath <-  "E:/new_downscaling/aoi/" # this must point to an existing directory with "Levy_MDV_actually.shp"
aoip <- list.files(aoipath, pattern="actually.shp", full.names = T)
clpath <- "E:/new_downscaling/coastline/Coastline_high_res_polygon/"

`%notin%` <- Negate(`%in%`)

## PATH TO TOOLS CMD

##### set paths for translating to SAGA
path_saga_norm <- "C:/OSGeo4W64/apps/saga-ltr/"
sagaCmd <- paste0(path_saga_norm, "saga_cmd.exe")
saga_outpath <- paste0(main, "SAGA_run/")

##### set paths for batch processing in HEG tool
# make an batchindir directory with the hdf files to batch convert and an batchoutdir directory with 
# the prm template file, where output files will be written to
# batchrunpath needs to be where MyHEG_batchScript.bat is located
batchrunpath <- "C:/Users/mleza/HEG/HEG_Win/bin/BatchRunning/BatchRunning/"
batchindir <- paste0(batchrunpath, "indir/") 
batchoutdir <- paste0(batchrunpath, "outdir/")
# for stacking images per month


source(paste0(scriptpath_organized, "0a_setup.R"))
#file.edit(paste0(scriptpath_organized, "0a_setup.R"))

# this is one raster with a complete coverage of the research area to use as a template 
template <- raster("E:/new_downscaling/clean_data/template_new.tif")

```

